WASP-4\lowercase{b} Arrived Early for the TESS Mission

### 2018-12-14 paper/ms.tex
grammar-hammer
paper: add phase-folded lightcurve figure

\section{New transits and system parameters}
\begin{figure}[t]
    \begin{center}
        \includegraphics[width=0.48\textwidth]{f1.png}
    \end{center}
    \vspace{-0.5cm}
    \caption{
        {\bf TESS transits of WASP-4b.} On the left, black points are
        TESS flux measurements, with a vertical offset applied. Blue
        curves are best-fit models. Rounded midtimes are printed in
        BTJD next to the appropriate transits.  The residuals are
        shown on the right.
        \label{fig:lightcurves}
    }
\end{figure}

@@

\begin{figure}[t]
    \begin{center}
        \leavevmode
        \includegraphics[width=0.48\textwidth]{f2.pdf}
    \end{center}
    \vspace{-0.6cm}
    \caption{
        {\bf TESS saw WASP-4b transit earlier than expected.} Both
        plots show observed transit times minus the pre-TESS ephemeris
        as a function of transit number.  Blue curves are the $\pm
        1\sigma$ credible interval around the constant-period
        prediction.  {\it Top:} The full timing dataset spans 10
        years.  Less precise points are more transparent.  {\it
        Bottom:} The binned TESS observation (red band) arrived $81.5
        \pm 11.8\ {\rm seconds}$ before it was expected.
        \label{fig:arrived_early}
    }
\end{figure}

@@

Having defined the free and fixed parameters, we then sampled over the
posterior using the algorithm proposed by
\citet{goodman_ensemble_2010} and implemented by
\citet{foreman-mackey_emcee_2013}.  Table~1 gives the results, which
are in reasonable agreement with the parameters reported by {\it
e.g.}, \citet{southworth_high-precision_2009} and
\citet{huitson_gemini_2017}.

@@

Figure~\ref{fig:lightcurves} shows the lightcurves, best-fit models,
and residuals.  Table~2 reports the mid-transit times and their
uncertainties.
Binning the residuals to 1-hour windows and taking the median absolute
deviation, the lightcurves have an RMS of $586\,{\rm ppm}\,{\rm
hr}^{1/2}$.  The predicted\footnote{\url{github.com/lgbouma/tnm}}
photon-counting noise is $410\,{\rm ppm}\,{\rm hr}^{1/2}$.  For a
$T\approx11.8$ star with a 9-pixel aperture, read noise is expected to
contribute an additional $202\,{\rm ppm}\,{\rm hr}^{1/2}$ in
quadrature.  The predicted zodiacal background contribution is
$673\,{\rm ppm}\,{\rm hr}^{1/2}$, based on the
\citet{winn_photonflux_2013} model, which \citet{Sullivan_2015} used
in planet detection simulations.  This zodiacal noise prediction seems
to have been an overestimate, given the observed RMS of WASP-4.  A
broader assessment of TESS's photometric performance is beyond
the scope of this work.

@@

\begin{figure*}[t]
    \begin{center}
        \leavevmode
        \includegraphics[width=0.9\textwidth]{f3.pdf}
    \end{center}
    \vspace{-0.5cm}
    \caption{
        {\bf Timing residuals and best-fit models for WASP-4b.}
        Points are observed transit times minus a linear fit to all
        observed times.  Less precise times are more transparent.  The
        best constant period model (gray line) is a poor fit to the
        observations.  The best-fit tidal decay model (blue) and
        apsidal precession model (orange) can fit the data.  The
        binned TESS point (red square) is for display purposes only;
        the individual TESS times are used to fit the models.
        \label{fig:times}
    }
\end{figure*}

@@

\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=0.417\textwidth]{f4.pdf}
  \end{center}
  \vspace{-0.5cm}
  \caption{
    {\bf WASP-4b should have a short eccentricity damping time and an average
    decay time compared to other hot Jupiters.  It seems to orbit a
    main sequence host.
    }
    {\bf \it Top:}
    Points are transiting giant planets from \citet{bonomo_gaps_2017},
    who measured the eccentricities using radial velocities.  Solid
    squares (including WASP-18b) have significant eccentricity
    detections.  The axes are chosen so that constant eccentricity
    damping timescales (Equation~\ref{eq:de_dt}) are lines.  The
    contours shown assume $Q_{\rm p}' = 10^5$.
    {\bf \it Middle:} 
    Axes are chosen so to that constant orbital decay timescales
    (Equation~\ref{eq:da_dt}) are lines.  The contours shown assume
    $Q_\star' = 10^7$.
    {\bf \it Bottom:}
    HR diagram highlighting WASP-12b, which shows timing variations
    similar to WASP-4b.  We also show hot Jupiters for which we
    detect no timing variation (see \S~\ref{sec:hj_verification}).
    \label{fig:context}
  }
\end{figure}

@@

If the planet's spin and orbit are synchronized and the star is slowly
rotating, then the semi-major axis and eccentricity evolve as
\citep[Appendix B of][]{metzger_optical_2012}
\begin{align}
  \frac{1}{\tau_{\rm e}} &=
  \frac{|\dot{e}|}{e} =
    \frac{63 \pi } {2 Q_{\rm p}' }
    \left( \frac{R_{\rm p}}{a} \right)^5
    \left( \frac{M_\star}{M_{\rm p}} \right)
    \left( \frac{1}{P} \right)
  \label{eq:de_dt}
  \\
  \frac{1}{\tau_{\rm a}} &=
  \frac{|\dot{a}|}{a} =
    \frac{9 \pi } {Q_\star' }
    \left( \frac{R_\star}{a} \right)^5
    \left( \frac{M_{\rm p}}{M_\star} \right)
    \left( \frac{1}{P} \right).
  \label{eq:da_dt}
\end{align}
The orbital period evolves as
\begin{equation}
\label{eq:dP_dt}
  \dot{P} = -\frac{27\pi}{2 Q_\star'}
            \left(\frac{M_{\rm p}}{M_\star}\right)
            \left(\frac{R_\star}{a}\right)^5.
\end{equation}
The modified quality factor of WASP-4 corresponding to the observed
value of $\dot{P}$ is then
\begin{equation}
	Q_\star' \approx 2.9\times10^4,
\end{equation}
which is rather small.  Jupiter has $Q_{\rm Jup}' \approx 1.4 \times
10^5$, based on the orbits of the Galilean moons
\citep{lainey_strong_2009}.  For stars, studies of the binary
eccentricity distribution have found $Q_\star' \approx 10^5 - 10^7$
\citep[{\it e.g.},][]{meibom_robust_2005,belczynski_compact_2008,
geller_direct_2013,milliman_wiyn_2014}.  The dissipation rates may
differ in hot Jupiter systems, since the tidal forcing frequencies and
amplitudes are different.  Observationally-driven studies of hot
Jupiter systems have found $Q_\star' \approx 10^5 - 10^8$, from quite
different models
\citep{jackson_observational_2009,hansen_calibration_2010,penev_constraining_2012,penev_empirical_2018,cameron_hierarchical_2018}.
For instance, motivated by the rapid rotation of some hot Jupiter
hosts \citep{pont_empirical_2009,penev_hats-18b_2016},
\citet{penev_empirical_2018} modeled the evolution of hot Jupiter
systems under the influence of a magnetized wind and a constant
phase-lag tide.  For WASP-4, their method gave $Q_\star' \approx
(1.2^{+1.0}_{-0.5})\times10^7$.  This strongly disagrees with the
quality factor we infer from the apparent period change.

@@

\begin{figure}[!t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.48\textwidth]{f5.pdf}
	\end{center}
  \vspace{-0.5cm}
	\caption{
		{\bf Further observations will be needed to
	    confirm and understand the timing variations of WASP-4b.}
		Dots are as in Figure~\ref{fig:times}.
		Lines are 100 random draws from the posteriors of the apsidal
		precession model (orange), and the orbital decay model (blue).    
		\label{fig:future}
	}
\end{figure}

@@

\begin{figure*}[ht!]
  \begin{center}
    \leavevmode
    \includegraphics[width=0.9\textwidth]{f6.pdf}
  \end{center}
  \vspace{-0.5cm}
  \caption{
    {\bf There is no evidence for a systematic offset between TESS
    times and the barycentric reference.} While the WASP-4b transits
    fell about 75 seconds earlier than expected, other well-observed
    hot Jupiters, in particular WASP-6b and WASP-18b, arrived on time.
    Ticks are observed TESS transit midtimes; the blue distribution is
    a kernel density estimate; the orange distribution is a gaussian
    centered on zero with standard deviation ($\sigma_{\rm
    predicted}$) calculated from the pre-TESS transit times.
    \label{fig:hjs}
  }
\end{figure*}

>>>>

\section{New transits and system parameters}

\begin{figure}[t]
    \begin{center}
        \includegraphics[width=0.48\textwidth]{f1.png}
    \end{center}
    \vspace{-0.5cm}
    \caption{
        {\bf Phase-folded transit of WASP-4b.} Black points are TESS
        flux measurements; yellow points are binned measurements.
        Using the fit to the phase-folded transit (blue line), we
        measure mid-transit times by fitting the individual transits
        (Figure~\ref{fig:lightcurves}).
        \label{fig:phasefold}
    }
\end{figure}

\begin{figure}[t]
    \begin{center}
        \includegraphics[width=0.48\textwidth]{f2.png}
    \end{center}
    \vspace{-0.5cm}
    \caption{
        {\bf Individual transits of WASP-4b.} On the left, black
        points are TESS flux measurements, with a vertical offset
        applied. Blue curves are best-fit models. Rounded midtimes are
        printed in BTJD next to the appropriate transits.  The
        residuals are shown on the right.
        \label{fig:lightcurves}
    }
\end{figure}

@@

\begin{figure}[t]
    \begin{center}
        \leavevmode
        \includegraphics[width=0.48\textwidth]{f3.pdf}
    \end{center}
    \vspace{-0.6cm}
    \caption{
        {\bf TESS saw WASP-4b transit earlier than expected.} Both
        plots show observed transit times minus the pre-TESS ephemeris
        as a function of transit number.  Blue curves are the $\pm
        1\sigma$ credible interval around the constant-period
        prediction.  {\it Top:} The full timing dataset spans 10
        years.  Less precise points are more transparent.  {\it
        Bottom:} The binned TESS observation (red band) arrived $81.5
        \pm 11.8\ {\rm seconds}$ before it was expected.
        \label{fig:arrived_early}
    }
\end{figure}

@@

Having defined the free and fixed parameters, we sampled over the
posterior using the algorithm proposed by
\citet{goodman_ensemble_2010} and implemented by
\citet{foreman-mackey_emcee_2013}.  Table~1 gives the results, which
are in reasonable agreement with the parameters reported by {\it
e.g.}, \citet{southworth_high-precision_2009} and
\citet{huitson_gemini_2017}.
Figure~\ref{fig:phasefold} shows the phase-folded lightcurve.

@@

Figure~\ref{fig:lightcurves} shows the lightcurve of each transit, the
best-fit models, and the residuals.  Table~2 reports the mid-transit
times and their uncertainties.
Binning the residuals to 1-hour windows and taking the median absolute
deviation, the lightcurves have an RMS of $586\,{\rm ppm}\,{\rm
hr}^{1/2}$.  The predicted\footnote{\url{github.com/lgbouma/tnm}}
photon-counting noise is $410\,{\rm ppm}\,{\rm hr}^{1/2}$.  For a
$T\approx11.8$ star with a 9-pixel aperture, read noise is expected to
contribute an additional $202\,{\rm ppm}\,{\rm hr}^{1/2}$ in
quadrature.  The predicted zodiacal background contribution is
$673\,{\rm ppm}\,{\rm hr}^{1/2}$, based on the
\citet{winn_photonflux_2013} model, which \citet{Sullivan_2015} used
in planet detection simulations.  This zodiacal noise prediction seems
to have been an overestimate, given the observed RMS of WASP-4.  A
broader assessment of TESS's photometric performance is beyond
the scope of this work.

@@

\begin{figure*}[t]
    \begin{center}
        \leavevmode
        \includegraphics[width=0.9\textwidth]{f4.pdf}
    \end{center}
    \vspace{-0.5cm}
    \caption{
        {\bf Timing residuals and best-fit models for WASP-4b.}
        Points are observed transit times minus a linear fit to all
        observed times.  Less precise times are more transparent.  The
        best constant period model (gray line) is a poor fit to the
        observations.  The best-fit tidal decay model (blue) and
        apsidal precession model (orange) can fit the data.  The
        binned TESS point (red square) is for display purposes only;
        the individual TESS times are used to fit the models.
        \label{fig:times}
    }
\end{figure*}

@@

\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=0.417\textwidth]{f5.pdf}
  \end{center}
  \vspace{-0.5cm}
  \caption{
    {\bf WASP-4b should have a short eccentricity damping time and an average
    decay time compared to other hot Jupiters.  It seems to orbit a
    main sequence host.
    }
    {\bf \it Top:}
    Points are transiting giant planets from \citet{bonomo_gaps_2017},
    who measured the eccentricities using radial velocities.  Solid
    squares (including WASP-18b) have significant eccentricity
    detections.  The axes are chosen so that constant eccentricity
    damping timescales (Equation~\ref{eq:de_dt}) are lines.  The
    contours shown assume $Q_{\rm p}' = 10^5$.
    {\bf \it Middle:} 
    Axes are chosen so to that constant orbital decay timescales
    (Equation~\ref{eq:da_dt}) are lines.  The contours shown assume
    $Q_\star' = 10^7$.
    {\bf \it Bottom:}
    HR diagram highlighting WASP-12b, which shows timing variations
    similar to WASP-4b.  We also show hot Jupiters for which we
    detect no timing variation (see \S~\ref{sec:hj_verification}).
    \label{fig:context}
  }
\end{figure}

@@

If the planet's spin and orbit are synchronized and the star is slowly
rotating, the semi-major axis and eccentricity then evolve as
\citep[Appendix B of][]{metzger_optical_2012}
\begin{align}
  \frac{1}{\tau_{\rm e}} &=
  \frac{|\dot{e}|}{e} =
    \frac{63 \pi } {2 Q_{\rm p}' }
    \left( \frac{R_{\rm p}}{a} \right)^5
    \left( \frac{M_\star}{M_{\rm p}} \right)
    \left( \frac{1}{P} \right)
  \label{eq:de_dt}
  \\
  \frac{1}{\tau_{\rm a}} &=
  \frac{|\dot{a}|}{a} =
    \frac{9 \pi } {Q_\star' }
    \left( \frac{R_\star}{a} \right)^5
    \left( \frac{M_{\rm p}}{M_\star} \right)
    \left( \frac{1}{P} \right).
  \label{eq:da_dt}
\end{align}
The orbital period evolves as
\begin{equation}
\label{eq:dP_dt}
  \dot{P} = -\frac{27\pi}{2 Q_\star'}
            \left(\frac{M_{\rm p}}{M_\star}\right)
            \left(\frac{R_\star}{a}\right)^5.
\end{equation}
The modified quality factor of WASP-4 corresponding to the observed
value of $\dot{P}$ is then
\begin{equation}
	Q_\star' \approx 2.9\times10^4,
\end{equation}
which is rather small.  Jupiter has $Q_{\rm Jup}' \approx 1.4 \times
10^5$, based on the orbits of the Galilean moons
\citep{lainey_strong_2009}.  For stars, studies of the binary
eccentricity distribution have found $Q_\star' \approx 10^5 - 10^7$
\citep[{\it e.g.},][]{meibom_robust_2005,belczynski_compact_2008,
geller_direct_2013,milliman_wiyn_2014}.  The dissipation rates may
differ in hot Jupiter systems, since the tidal forcing frequencies and
amplitudes are different.  Observationally-driven studies of hot
Jupiter systems have found $Q_\star' \approx 10^5 - 10^8$, from quite
different models
\citep{jackson_observational_2009,hansen_calibration_2010,penev_constraining_2012,penev_empirical_2018,cameron_hierarchical_2018}.
For instance, motivated by the rapid rotation of some hot Jupiter
hosts \citep{pont_empirical_2009,penev_hats-18b_2016},
\citet{penev_empirical_2018} modeled the evolution of hot Jupiter
systems under the influence of a magnetized wind and a constant
phase-lag tide.  For WASP-4, their method gave $Q_\star' \approx
(1.2^{+1.0}_{-0.5})\times10^7$.  This strongly disagrees with the
quality factor we infer from the apparent period change.

@@

\begin{figure}[!t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.48\textwidth]{f6.pdf}
	\end{center}
  \vspace{-0.5cm}
	\caption{
		{\bf Further observations will be needed to
	    confirm and understand the timing variations of WASP-4b.}
		Dots are as in Figure~\ref{fig:times}.
		Lines are 100 random draws from the posteriors of the apsidal
		precession model (orange), and the orbital decay model (blue).    
		\label{fig:future}
	}
\end{figure}

@@

\begin{figure*}[ht!]
  \begin{center}
    \leavevmode
    \includegraphics[width=0.9\textwidth]{f7.pdf}
  \end{center}
  \vspace{-0.5cm}
  \caption{
    {\bf There is no evidence for a systematic offset between TESS
    times and the barycentric reference.} While the WASP-4b transits
    fell about 75 seconds earlier than expected, other well-observed
    hot Jupiters, in particular WASP-6b and WASP-18b, arrived on time.
    Ticks are observed TESS transit midtimes; the blue distribution is
    a kernel density estimate; the orange distribution is a gaussian
    centered on zero with standard deviation ($\sigma_{\rm
    predicted}$) calculated from the pre-TESS transit times.
    \label{fig:hjs}
  }
\end{figure*}
